---
title: <font size = '15'>Loan Defaulting Investigation</font>
subtitle: <font size ='8'>For Bank Clients In India</font>
author: <font size ='6'>Amigos <br> Orestis Patsalides, Nicolas Georgiades and Iskandar Sawalha</font>
institute: <font size ='8'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
    transition: convex
    theme: league
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
Loans <- read.csv("/cloud/project/data/Loans.csv")
```

```{r test, include=FALSE, message=FALSE}
Loans %>%
  ggplot() +
  geom_hex(aes(x = Experience, y = log10(Income))) +
  facet_wrap(~ risk_flag) +
  ggtitle("Experience vs Income", subtitle = "By Loan Defaulting Status") +
  ylab("Income (log10)")
```

```{r data-split, include=FALSE, message=FALSE}
Loans_2 <- Loans %>%
   mutate(risk_flag = factor(risk_flag, labels = c("No", "Yes")))

med_inc <- median(Loans_2$Income)
high_inc <- (Loans_2$Income) >= med_inc

Loans_3 <- Loans_2 %>%
  filter(Income >= med_inc)

set.seed(1122)
Loans_split <- initial_split(Loans_3, prop = 0.80)

train_data <- training(Loans_split)
test_data <- testing(Loans_split)
```

# <font size = '12'> Introducing our Investigation </font> {data-background-image="img/s2.png"}

How does a person’s situation and current status affect Loan Defaulting?

We will try to assess the above question by examining how Experience, Income and Marital Status affect Loan Defaulting, for bank clients in India.

We will do this by the use of statistical models and graphs.

# <font size = '12'> Introducing our Data </font> {data-background-image="img/s3.png"}

Our data frame shows observations gathered for different individuals in India
related to their current status in life and whether they were able to
repay a loan or not. All values were provided at the time of the loan application at several banks India.

## <font size = '9'> Here is a glimpse of our data frame </font> {data-background-image="img/s3.png"}

```{r Loans-table, echo = FALSE}
kable(head(Loans_3[1:6]), format = "html")
```

*Source: kaggle.com*

# <font size = '12'> Does Income really Increase with Experience </font>

```{r initial-relationship, echo=FALSE, message=FALSE}
train_data %>%
  group_by(Experience, Married.Single) %>%
  summarise("med_inc" = median(Income)) %>%
  ggplot(aes(x = Experience, y = med_inc)) +
  geom_point() +
  facet_wrap(~ Married.Single) +
  geom_smooth(aes(color = "red")) + 
  scale_y_continuous(labels = label_dollar(scale = 1/100000, 
                                           prefix = "£", 
                                           suffix = "K", 
                                           accuracy = 1)) +
  labs(title = "Median Income per Year of Experience",
       subtitle = "By Marital Status",
       x = "Experience (years)",
       y = "Median Income (£)") +
  theme_minimal() + 
  theme(legend.position = "none")
```

## <font size = "9"> Not Always </font>

<font size = "5" align = "left">

From the graph we can see that for married people income generally decreases with experience, whilst for single people it increases. 

Focusing on married people, we can see that income starts higher than for single people, as they have 2 salaries. There is a steep decrease which can be explained by the fact that possibly the woman stops working after the marriage, as is usually the trend in developing countries with less developed societies. Then, after 7 years of experience, the income starts to rise, due to the fact that the people have more experience, thus are more productive and are rewarded for this. However, after 12 years income starts decreasing again, likely because after a certain age people become fatigued of their job and their productivity decreases and some even retire (a number which increases as the years of experience increase), leading to the effect observed. 

For single people, the general trend for the first 15 years is for income to grow with experience. Although, after 15 years people grow bored of their jobs, meaning they become less productive or consider retiring, both results increasing as the years go by, thus a decrease in income is observed.

</font>

# <font size = '12'> How does Income affect Loan Defaulting </font>

```{r relationship1, echo=FALSE, message=FALSE}
train_data %>%
filter((Income) >= med_inc) %>%
  ggplot(aes(x = Income, fill = risk_flag)) +
  geom_histogram(binwidth = 200000, show.legend = FALSE) +
  facet_wrap(~ risk_flag) +
  scale_x_continuous(breaks = seq(med_inc, 10000000, by = 2000000),
                     labels = label_dollar(scale = 1/100000,
                                           prefix = "£",
                                           suffix = "K",
                                           accuracy = 1)) +
  labs(title = "High Income Distribution",
       subtitle = "By Loan Defaulting",
       x = "High Income (£)",
       y = "Frequency") +
  scale_fill_manual(values = c("No" = "Red", "Yes" = "Green")) +
  theme_minimal()
```

```{r logistic-regression1, include=FALSE, message=FALSE}
Loans_Inc_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(risk_flag ~ Income, Loans_3, family = "binomial")
tidy(Loans_Inc_fit)
```

## <font size = "9"> TITLE </font>

<font size = "5" align = "left">

In order to be able to investigate the above question we needed to filter our data so that it only consisted of clients earning an income which was higher than the median income. Clients who fell in this category were considered to be earning a high income.

From the graph we can clearly see that people earning high income are less likely to default on a loan, since the frequency of those who were able to repay their loan is substantially higher than the one representing clients who failed to repay.

As a result, it can be concluded that a person earning a high income (higher than the median income of all clients, for the purpose of this investigation) does affect at a high extent if a client will be able to repay a loan or not.

</font>

# <font size = '12'> How does Marriage Status affect this relationship </font>

```{r relationship2, echo=FALSE, message=FALSE}
train_data %>%
filter((Income) >= med_inc) %>%
  ggplot(aes(x = Income, fill = risk_flag)) +
  geom_histogram(binwidth = 200000, show.legend = FALSE) +
  facet_wrap(Married.Single ~ risk_flag) +
  scale_x_continuous(breaks = seq(med_inc, 10000000, by = 2000000),
                     labels = label_dollar(scale = 1/100000, 
                                           prefix = "£", 
                                           suffix = "K",
                                           accuracy = 1)) +
  labs(title = "High Income Distribution",
       subtitle = "By Loan Defaulting and Marital Status",
       x = "High Income (£)",
       y = "Frequency") +
  scale_fill_manual(values = c("No" = "Red", "Yes" = "Green")) +
  theme_minimal()
```

```{r logistic-regression2, include=FALSE, message=FALSE}
Loans_Inc_Mar_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(risk_flag ~ Income + Married.Single, Loans_3, family = "binomial")
tidy(Loans_Inc_Mar_fit)
```

## <font size = "9"> Slightly </font>

<font size = "5" align = "left">

Generally, from the above faceted graph, we can observe that the number clients who default their loan is greater than that of clients who don't default despite their marriage status. Furthermore, the number of single clients is greater than that of married clients for both defaulted and non-defaulted loans.

Paying greater attention, we can see that married people are less likely to default on a loan. The relative height difference of the bars for married clients who didn't repay their loan compared to those who did is smaller that the equivalent difference for single clients. However the difference between clients with different marriage statuses does not appear to be very large.

As a result, it can be concluded that if a client (with a higher than the median income of all clients) is married does affect  if he/she will be able to repay a loan or not, but only at a small extent.

</font>

# <font size = '12'> How Accurate is the Model in Predicting future Loan Defaulting </font>
*<font size = '6'> Based on High Income and Marital Status </font>*

```{r prediction, echo=FALSE, message=FALSE}
Loans_rec <- recipe(
  risk_flag ~ Income + Married.Single,
  data = train_data
)

Loans_mod <- logistic_reg() %>%
  set_engine("glm")

Loans_wflow <- workflow() %>%
  add_model(Loans_mod) %>%
  add_recipe(Loans_rec)

Loans_fit <- Loans_wflow %>%
  fit(data = train_data)

Loans_pred <- predict(Loans_fit,
                      test_data,
                      type = "prob") %>%
  bind_cols(test_data %>% select(risk_flag, Income, Married.Single))

Loans_pred %>%
  arrange(desc(.pred_Yes)) %>%
  print(n = 10)
```

## <font size = '9'> Graphical Representation of the Model's Accuracy </font>

```{r roc-curve, echo=FALSE, message=FALSE}
Loans_pred %>%
  roc_curve(truth = risk_flag,
            .pred_Yes,
            event_level = "second") %>%
  autoplot()

train_data %>%
  count(risk_flag)

Loans_pred %>%
  mutate("Prediction" = ifelse(test = .pred_Yes > 0.50, 1, 0)) %>%
  count(Prediction)
```

# Layouts

You can use plain text

- or bullet points

<div>
  <div style="float: left; width: 50%;">or text in two columns</div>
  <div style="float: left; width: 50%;">like this</div>
</div>

because you can combine HTML and Markdown syntax.

<p style="font-size: 8pt; position: absolute; bottom: -5em; width: 100%; text-align: center;">*You can also add footnotes</p>

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

## Plot and text

::: columns
:::: column
* Some text
* goes here
::::

:::: column
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

## Different widths

::: columns
:::: {.column width="40%"}
* Some text
* goes here
::::

:::: {.column width="60%"}
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

## {data-background-image="https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80"}

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs (\$...\$), e.g. $\alpha+\beta$. You can use the display style with double dollar signs:

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

# Feeling adventurous?

You can learn more about the options available at the following pages:

- https://github.com/rstudio/revealjs

- https://bookdown.org/yihui/rmarkdown/revealjs.html

- http://www.jenniferbradham.org/post/reveal_js_markdown/

- https://revealjs.com/