---
title: "Loan Defaulting Investigation"
subtitle: <font size ='7'>For Bank Clients In India</font>
author: <font size ='6'>Amigos <br> Orestis Patsalides, Nicolas Georgiades and Iskandar Sawalha </font>
institute: <font size ='10'>University of Edinburgh</font>
date: <font size = '6'>`r Sys.Date()`</font>
output:
  revealjs::revealjs_presentation:
    self_contained: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(rmarkdown)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(revealjs)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
Loans <- read.csv("/cloud/project/data/Loans.csv")
```

## Introducing our Investigation  {data-background-image="img/s2.png"}


How does a person’s situation and current status affect Loan Defaulting?

We will try to assess the above question by examining how Age, Income and Marital Status affect Loan Defaulting, for bank clients in India.

We will do this by the use of statistical models and graphs.

# Introducing our Data {data-background-image="img/s3.png"}



Our data frame shows observations gathered for different individuals in India
related to their current status in life and whether they were able to
repay a loan or not. All values were provided at the time of the loan application at several banks India.

----

Here is a glimpse of our data frame.

```{r Loans-table, echo = FALSE}
kable(head(Loans[1:6]), format = "html")
```

```{r,echo=FALSE, message=FALSE, echo=FALSE}
Loans_2 <- Loans %>%
   mutate(risk_flag = factor(risk_flag, labels = c("Did not Default on Loan", "Defaulted on Loan")))

med_inc <- median(Loans_2$Income)
high_inc <- (Loans_2$Income) >= med_inc

set.seed(1122)
Loans_split <- initial_split(Loans_2, prop = 0.80)

train_data <- training(Loans_split)
test_data <- testing(Loans_split)

Loans_3 <- Loans_2 %>%
  filter(Income >= med_inc)
```

Source: kaggle.com

# Does Income really Increse with Experience?

```{r,echo=FALSE, message=FALSE, echo=FALSE}
train_data %>%
  group_by(Experience, Married.Single) %>%
  summarise("med_inc" = median(Income)) %>%
  ggplot(aes(x= Experience, y = med_inc)) +
  geom_point() +
  facet_wrap(~ Married.Single) +
  geom_smooth()+
  scale_y_continuous(labels = label_dollar(scale = 1/100,
                                           prefix = "£"))+
  labs(
    title = "Experience of person in years vs. median income of person in Euros",
    x = "Experience in Years",
    y = "Median Income(Euros)",
    subtitle = "Faceted by Wether a Person is Married or Single"
  ) +
  theme_minimal()
```

----

*Not Always*

<font size="5" align="left">

From the graph we can see that for married people income generally decreases with experience, whilst for single people it increases. 

Focusing on married people, we can see that income starts higher than for single people, as they have 2 salaries. There is a steep decrease which can be explained by the fact that possibly the woman stops working after the marriage, as is usually the trend in developing countries with less developed societies. Then, after 7 years of experience, the income starts to rise, due to the fact that the people have more experience, thus are more productive and are rewarded for this. However, after 12 years income starts decreasing again, likely because after a certain age people become fatigued of their job and their productivity decreases and some even retire (a number which increases as the years of experience increase), leading to the effect observed. 

For single people, the general trend for the first 15 years is for income to grow with experience. Although, after 15 years people grow bored of their jobs, meaning they become less productive or consider retiring, both results increasing as the years go by, thus a decrease in income is observed.

</font>

# How does Income affect Loan Defaulting

```{r,echo=FALSE, message=FALSE}
train_data %>%
filter((Income) >= med_inc) %>%
  ggplot(aes(x = Income, fill = risk_flag)) +
  geom_histogram(binwidth = 200000, show.legend = FALSE) +
  facet_wrap(~ risk_flag) +
  scale_x_continuous(breaks = seq(med_inc, 10000000, by = 2000000),
                     labels = label_dollar(scale = 1/100000,
                                           prefix = "£",
                                           suffix = "K",
                                           accuracy = 1)) +
  labs(
    title = "Income of People with an Income Greater than the Median income vs Amount of People",
    subtitle = "Faceted by Wether People Defaulted on their Loans or not",
    x = "Income(Euros)",
    y = "Number of People",
  ) +
  theme_minimal()

  

Loans_Inc_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(risk_flag ~ Income, Loans_2, family = "binomial")
tidy(Loans_Inc_fit)
```

----

*How does Marriage Status affect this relationship*

```{r,echo=FALSE, message=FALSE}
train_data %>%
filter((Income) >= med_inc) %>%
  ggplot(aes(x = Income, colour = risk_flag, fill = risk_flag)) +
  geom_histogram(binwidth = 200000, show.legend = FALSE) +
  facet_wrap(Married.Single ~ risk_flag) +
  scale_x_continuous(breaks = seq(med_inc, 10000000, by = 2000000),
                     labels = label_dollar(scale = 1/100000, 
                                           prefix = "£", 
                                           suffix = "K",
                                           accuracy = 1)) +
  labs(
    title = "Income of People with an Income Greater than the Median income vs Amount of People",
    subtitle = "Faceted by Wether People Defaulted on their Loans or not and Wether they are Single or Married",
    x = "Income(Euros)",
    y = "Number of People",
  ) +
  theme_minimal()

Loans_Inc_Mar_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(risk_flag ~  + Married.Single, Loans)
tidy(Loans_Inc_Mar_fit)
```

----

```{r,echo=FALSE, message=FALSE}
Loans_rec <- recipe(
  risk_flag ~ 
)
```


# Layouts

You can use plain text

- or bullet points

<div>
  <div style="float: left; width: 50%;">or text in two columns</div>
  <div style="float: left; width: 50%;">like this</div>
</div>

because you can combine HTML and Markdown syntax.

<p style="font-size: 8pt; position: absolute; bottom: -5em; width: 100%; text-align: center;">*You can also add footnotes</p>

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

## Plot and text

::: columns
:::: column
* Some text
* goes here
::::

:::: column
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

## Different widths

::: columns
:::: {.column width="40%"}
* Some text
* goes here
::::

:::: {.column width="60%"}
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
::::
:::

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

## {data-background-image="https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80"}

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs (\$...\$), e.g. $\alpha+\beta$. You can use the display style with double dollar signs:

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

# Feeling adventurous?

You can learn more about the options available at the following pages:

- https://github.com/rstudio/revealjs

- https://bookdown.org/yihui/rmarkdown/revealjs.html

- http://www.jenniferbradham.org/post/reveal_js_markdown/

- https://revealjs.com/